Versatile Robust Clustering of Ad Hoc Cognitive Radio Network

### 2016-11-30 journal-draft/bare_jrnl_comsoc.tex
modified system model

We consider a set of cognitive radio users $\mathcal{N}$ and a set of primary users distributed over a given area.
The CR users use a set of licensed channels $\mathcal{K}$ opportunistically. 
The CR users are allowed to transmit on any channel $k \in \mathcal{K}$ only if no primary user is detected on channel $k$. 
CR users conduct spectrum sensing independently and sequentially on all licensed channels.
We assume that every node can detect the presence of an active primary user on each channel with certain accuracy. \footnote{The spectrum availability can be validated with a certain probability of detection. Spectrum sensing/validation is out of the scope of this paper.}
We adopt the unit disk model~\cite{unitDiskModel} for the transmission of both primary and CR users.
As to the relation between the primary and secondary users, If a CR node $i$ locates within the transmission range of an active primary user $p$, then $i$ is not allowed to use the channel which is being used by $p$.
On the other hand, If $i$ is not in the transmission range of primary user $p$, $i$ can certainly not detect the presence of $p$.
As the result of spectrum sensing, $K_i \subseteq \mathcal{K}$ denotes the set of licensed channels which can be used by $i$.
As the transmission range of primary users is limited and secondary users have different locations, different secondary users may have different views of the spectrum availability, i.e., for any $i, j \in \mathcal{N}$, $K_i = K_{j}$ does not necessarily hold.	
Among the secondary users, we denote $\text{Nb}(i)$ as user $i$'s neighborhood, which consists of the CR nodes located within the transmission range of $i$.

We assume there is one dedicated control channel which is used to exchange signaling messages during clustering process.
This control channel could be one of the ISM bands or other reserved spectrum which is exclusively used for transmitting control messages.
\footnote{Actually, the control messages involved in the clustering process can also be transmitted on the available licensed channels through a rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}, i.e., two neighboring nodes establish communication on the same channel.}
Over the control channel, a secondary user $i$ can exchange its spectrum sensing result $K_i$ to all its one hop neighbors $\text{Nb}(i)$.
In the following, we refer to the licensed channel as \textit{channel} in general, unless we mention the control channel explicitly. 
Then, a cognitive radio network can be represented as a graph $G = (\mathcal{N}, E)$, where $E \subseteq \mathcal{N} \times \mathcal{N}$ such that $\{i, j\} \in E$  if, and only if,$K_{i} \cap K_{j}\neq \emptyset$ and $d_{i,j} < r$ where $d_{i,j}$ is the distance between $i, j$ and $r$ is the radius of the secondary user's transmission range.

A cluster $C$ is a set of secondary nodes, which consists a cluster head $h_C$ and a number of cluster members.
The cluster head is able to communicate with any cluster member directly.
In other terms, for any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
In this paper, we denote a cluster as $C$, or $C(i)$ when we want to point out the cluster head is $i$.
Cluster size of $C(i)$ is written as $|C(i)|$.
$ K(C) = \cap_{i\in C} K_i$, $K(C)$ denotes the set of common licensed channels in cluster $C$.
Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model and the following problem description are listed in the Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & collection of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h_C$ & the cluster head of a cluster C\\
$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
$C_i$  & the $i$th potential cluster in a collection of clusters \\
 $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
\bottomrule
\end{tabular}
\end{table}

@@

As introduced in Section I, in order to be robust against primary users' activity, the formed clusters should have more CCs to expect longer time expectancy.
Meanwhile, the sizes of the formed clusters should not diverge from the desired size greatly.
The formation of small clusters or the \textit{singleton clusters}, \ie the cluster which has only one CR node, contradicts the motivation of forming clusters, as the benefits brought in by the collective of the cluster members are compromised.
On the other hand, large clusters are not preferred in some scenarios neither, \eg for the CRN composed with resource limited users, managing the cluster members in a large cluster is a substantial burden. 
Hence, the cluster size should fall in a desired range according to different application scenarios~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.
Considering the above mentioned requirements, we present the definition of robust clustering problem as follows.

>>>>

We consider a set of CR users $\mathcal{N}$ and a set of primary users distributed over a given area.
A set of licensed channels $\mathcal{K}$ is available for the primary users. 
The CR users are allowed to transmit on channel $k \in \mathcal{K}$ only if no primary user is detected on channel $k$. 
CR users conduct spectrum sensing independently and sequentially on all licensed channels.\footnote{We assume that every node can detect the presence of an active primary user on each channel with certain accuracy. The spectrum availability can be validated with a certain probability of detection. Spectrum sensing/validation is out of the scope of this paper.}
We adopt the unit disk model~\cite{unitDiskModel} for both primary and CR users' transmission.
%As to the relation between the primary and secondary users,
If a CR node $i$ locates within the transmission range of an active primary user $p$, then $i$ is not allowed to use the channel which is being used by $p$.
%On the other hand, if $i$ is not in the transmission range of primary user $p$, $i$ can certainly not detect the presence of $p$.
We assume the primary users change their operation channels slowly, then we omit the time index for the spectrum availability, \ie as the result of spectrum sensing, $K_i \subseteq \mathcal{K}$ denotes the set of available licensed channels for CR user $i$ at a time point.
As the transmission range of primary users is limited and secondary users have different locations, different secondary users may have different views of the spectrum availability, i.e., for any $i, j \in \mathcal{N}$, $K_i = K_{j}$ does not necessarily hold.
A cognitive radio network can be represented as a graph $G = (\mathcal{N}, E)$, where $E \subseteq \mathcal{N} \times \mathcal{N}$ such that $\{i, j\} \in E$ if and only if $K_{i} \cap K_{j}\neq \emptyset$ and $d_{i,j} < r$, where $d_{i,j}$ is the distance between $i, j$, $r$ is the radius of secondary user's transmission range. 
%
Among the secondary users, we denote $\text{Nb}(i)$ as user $i$'s neighborhood, which consists of the CR nodes located within the transmission range of $i$.

We assume there is one dedicated control channel which is used to exchange signaling messages during clustering process.
This control channel could be one of the ISM bands or other reserved spectrum which is exclusively used for transmitting control messages.\footnote{Actually, the control messages involved in the clustering process can also be transmitted on the available licensed channels through a rendezvous process by channel hopping~\cite{channelHopping_Rendezvous_2014, Gu_distributed_rendezvous_2014}, i.e., two neighboring nodes establish communication on the same channel.}
Over the control channel, a secondary user $i$ can exchange its spectrum sensing result $K_i$ with all its one hop neighbors $\text{Nb}(i)$.
In the following, we refer to the licensed channel as \textit{channel} in general, and will explicitly mention the dedicated control channel if necessary.

A cluster $C$ is a set of secondary nodes, which consists of a cluster head $h_C$ and a number of cluster members.
The cluster head is able to communicate with any cluster member directly.
$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
Cluster size of $C$ is written as $|C|$.
Cluster $C(i)$ means the cluster head of this cluster is $i$.
$ K(C) = \cap_{i\in C} K_i$, $K(C)$, which denotes the set of CCs in cluster $C$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model and the following problem description are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & collection of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h_C$ & the cluster head of a cluster C\\
$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
$C_i$  & the $i$th potential cluster in a collection of clusters \\
 $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
\bottomrule
\end{tabular}
\end{table}

@@

As introduced in Section~\ref{intro}, in order to be robust against primary users' activity, the formed clusters should have more CCs to expect longer time expectancy.
Meanwhile, the sizes of the formed clusters should not diverge from the desired size greatly.
The formation of small clusters or the \textit{singleton clusters}, \ie the cluster which has only one CR node, contradicts the motivation of forming clusters, as the benefits brought in by the collective of the cluster members are compromised.
On the other hand, large clusters are not preferred in some scenarios neither, \eg for the CRN composed with resource limited users, managing the cluster members in a large cluster is a substantial burden. 
Hence, the cluster size should fall in a desired range according to different application scenarios~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.
Considering the above mentioned requirements, we present the definition of robust clustering problem as follows.
